### 2019-10-07 main_text/main_text.tex
Edits to intro

\section*{Introduction}
Genetic regulatory circuits are fundamental building blocks of functioning
cells and organisms. One abundant class of these circuits are genetic
switches. They can be constructed in various ways and have different
functions. What all of them have in common is that the system is bistable:
there are two stable steady-states that it can be in. The distribution of
gene expression in a cell culture can then be bimodal. This is not to be
confused with mere stochastic bimodality, where the system is not stable,
but the gene expression in each cell is fluctuating between the two levels.

One classic example of a genetic switch is a system where two repressor
proteins each regulate the transcription of the
other~\cite{Jacob1961,Gardner2000} (illustrated schematically in
Fig~\ref{fig1:switches}). Here, the two stable states are high expression of
the first protein and low expression of the second one or the opposite. This
switch enables the system to have a memory: if something induces expression
of either one of the proteins, the system will remain in this state until a
significant perturbation occurs. Another well known and even simpler example
is an auto-activating circuit in which a protein activates its own
transcription~\cite{Wolf1998}. This gives the system an ``on-off''-switch.
%\todo{make eps}

\begin{figure}%[h!]
	\centering
	\includegraphics{media/Circuits.eps}
	\caption{{\bf A schematic of different genetic switches.}
	(A) and (B) show the two most well-known genetic switches that are
	mentioned in the text. In (C), a very much simplified version of the
	circuit that we investigate in this paper can be seen, where the
	similarity to the switch in (B) becomes obvious. A complete version of
	the model circuit can be found in Fig~\ref{fig2:model}.}
	\label{fig1:switches}
\end{figure}

@@

For our modeling, we focus on the \emph{xapABR} genetic circuit from
\emph{E. coli} as a case study. It is similar to \emph{lac}, but less
complex. Instead of lactose, its purpose is to make use of the nucleoside
xanthosine as an energy source~\cite{Buxton1980,Hammer-Jespersen1980}. The
circuit is made up of two genes: one that encodes for XapR and another one
that encodes for XapA and XapB. XapR appears to be a transcription factor
that is induced by xanthosine and activates the \emph{xapAB} promoter. It
seems to have two binding sites on the promoter, but the promoter
architecture and function is not fully understood up to now. The
transcription of \emph{xapR} seems to be constitutive and not
auto-regulated~\cite{Seeger1995}. Structural homology to other transcription
factors suggests that XapR appears in dimers where one dimer can bind two
xanthosine molecules~\cite{Joergensen1999}. The protein XapA is a purine
nucleoside phosphorylase that degrades xanthosine into components (ribose
and xanthine) that can be fed into metabolic
pathways~\cite{Buxton1980,Hammer-Jespersen1980}. XapB on the other hand is a
membrane transporter of xanthosine~\cite{Seeger1995,Norholm2001}.

>>>>

\section*{Introduction}
Genetic regulatory circuits are fundamental building blocks of functioning
cells and organisms. One abundant class of these circuits are genetic
switches. Although their construction and function may differ,
their common feature is bistability: their output gene expression
will flow to and remain at one of two steady-state levels.
The distribution of
gene expression in a cell culture can then be bimodal. This is not to be
confused with mere stochastic bimodality, where the system is not stable,
and the gene expression in each cell can fluctuate between the two levels.

One classic example of a genetic switch is a system where two repressor
proteins each regulate the transcription of the
other~\cite{Jacob1961,Gardner2000} (illustrated schematically in
Fig~\ref{fig1:switches}). Here, one stable state is high expression of
the first protein and low expression of the second,
and the second stable state is the opposite. This
switch enables the system to have a memory: if something induces expression
of either one of the proteins, the system will remain in this state until a
significant perturbation occurs. Another well known and even simpler example
is an auto-activating circuit in which a protein activates its own
transcription~\cite{Wolf1998}. This gives the system an ``on-off'' switch.

\begin{figure}%[h!]
	\centering
	\includegraphics{media/Circuits.eps}
	\caption{{\bf A schematic of different genetic switches.}
	(A) and (B) show the two most well-known genetic switches:
	(A) two mutual repressors and (B) a self-activating gene.
	In (C), a very much simplified version of the
	circuit that we investigate in this paper can be seen, where the
	similarity to the switch in (B) becomes obvious. A complete version of
	the model circuit can be found in Fig~\ref{fig2:model}.}
	\label{fig1:switches}
\end{figure}

@@

\textit{MM: I like the above paragraph, but we should also add something
about the TF \& frame it as resourse utility: xapR, lacI, and I suspect
others of these TFs happen to be constitutively expressed at low copy
number. They act as ``always on'' sensors that trigger a switch to produce
massive numbers of metabolic enzymes only when their substrate is
available.}

For our modeling, we focus on the \emph{xapABR} genetic circuit from
\emph{E. coli} as a case study. It is similar to \emph{lac}, but less
complex. Instead of lactose, its purpose is to make use of the nucleoside
xanthosine as an energy source~\cite{Buxton1980,Hammer-Jespersen1980}. The
circuit is made up of two operons: one that encodes for XapR and another
that encodes for XapA and XapB. XapR appears to be a transcription factor
that is induced by xanthosine and activates the \emph{xapAB} promoter.
This promoter seems to have two binding sites for XapR, but the promoter
architecture and function is not yet fully understood. The
transcription of \emph{xapR} seems to be constitutive and not
auto-regulated~\cite{Seeger1995}. Structural homology to other transcription
factors suggests that XapR appears in dimers where one dimer can bind two
xanthosine molecules~\cite{Joergensen1999}. The protein XapA is a purine
nucleoside phosphorylase that degrades xanthosine into components (ribose
and xanthine) that can be fed into metabolic
pathways~\cite{Buxton1980,Hammer-Jespersen1980}. XapB on the other hand is a
membrane transporter of xanthosine~\cite{Seeger1995,Norholm2001}.
